---
title: "Assignment 3.2 - Student Assignment Performance Data"
author: "Demond Love"
date: "6/24/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results = FALSE}
library(readr)
library(ggplot2)
library(pastecs)

path = file.path('/Users/Love/Documents/DSC 520 Statistics for Data Science/Week 3 DSC 520/DSC 520 Week 3/Exams.dat')
file = read_tsv(path, col_names = TRUE)
SELECT = subset(file, select = c(exam, computer, lectures, numeracy, stats, uni))

```

Grab statistics for each variable.

```{r}
stat.desc(SELECT)
```

The first three variables in the data are a percentage based on 100 points.

```{r}
qplot(SELECT$exam)
```

This distribution look bimodal.

```{r}
qplot(SELECT$computer)
```

This looks semi-normal; slight negative skew, but probably isn't statistically significant.

```{r}
qplot(SELECT$lectures)
```

This also looks semi-normal; slight negative skew, but probably isn't statistically significant.

The fourth variable is based on a score achieved out of 15 points.

```{r}
qplot(SELECT$numeracy)
```

This looks significantly positively skewed.

The stats variable are the letter grades for the Stats exam where A=1, B=2, C=3, D=4, F=5. 
```{r}
qplot(SELECT$stats)
```

This looks negatively skewed, since the mode value is a 5.

Doesn't really make sense to plot uni, since the uni variable indicates which professor taught that section of the course. 0=Professor George, 1=Professor Jeff, and 3=Professor Tushmann.

Next, compare variables for normality normality numerically.

```{r}
round(stat.desc(SELECT, basic = FALSE, norm = TRUE), digits = 3)
```


Comparing each professor against the grades in each exam.

```{r}
univexam = ggplot(SELECT, aes(x = uni, y = exam))
univexam + geom_point() + geom_smooth()
```

Here, you can tell that Professor Jeff has significantlly higher scores on the first exam.

```{r}
univcomputer = ggplot(SELECT, aes(x = uni, y = computer))
univcomputer + geom_smooth() + geom_point()
```

Although Professor George and Professor Jeff have simialr scores, you can see that Professor Tushmann's scores are lower on the second exam.

```{r}
univlectures = ggplot(SELECT, aes(x = uni, y = lectures))
univlectures + geom_smooth() + geom_point()
```

Here, all three professor's students faired almost the same.

Next, transforming the numeracy variable to correct it's positive skew.

```{r}
lognumeracy = log(SELECT$numeracy)
qplot(lognumeracy)
round(stat.desc(lognumeracy, basic = FALSE, norm = TRUE), digits = 3)
```

Looks like the log transformation seems to have worked, since skew.2SE decreased from 2.376 to -1.103. Although it is still showing statistically significant skewness, the variable is significantly more normal.

And finally, we will run the Shapiro-Wilk test, and plot each variable accordingly.

```{r}
shapiro.test(SELECT$exam)
qplot(SELECT$exam, stat = 'qq')
shapiro.test(SELECT$computer)
qplot(SELECT$computer, stat = 'qq')
shapiro.test(SELECT$lectures)
qplot(SELECT$lectures, stat = 'qq')
shapiro.test(SELECT$numeracy)
qplot(SELECT$numeracy, stat = 'qq')
shapiro.test(SELECT$stats)
qplot(SELECT$stats, stat = 'qq')
```

