---
title: "DSC 520 Assignment 2.3"
author: "Demond Love"
date: "6/17/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**There are several different ways to import text files (.csv is a common text file) in R as you have learned. Use one of those import methods to import your dataset for this assignment. Remember to make sure you setup a Project and have set your working directory as needed (hint: setwd(“…..”).Choose one of the Weather datasets [AKWeather or NEWeather] to use for this activity.**

**What categories/variables of data do you have in your Weather dataset?**
```{r}
akweather = read.csv("akweather.csv", stringsAsFactors = FALSE)
akweather
```

**Use the following R functions: View(); str(); nrow(); ncol() to investigate the dataset you are working with. Report the output from each function and provide a brief explanation of the output. Remember the following resources, outside of R Studio can be used for additional help: Google – Just add R to the end of your search criteria R Bloggers - https://www.r-bloggers.com/**
```{r}
# View(akweather)
```
This function produces the equivalent of an Excel Spreadsheet of your dataset. It is maybe my favorite feature so far when comparing it to Python. Coming from SQL, seeing the data in a true tabular format is much more intuitive for me.
```{r}
str(akweather)
```
This function shows the structure of a dataset. Which will display the class, the number of rows, the number of columns, the class of the values in the rows, etc.

```{r}
nrow(akweather)
```
The nrow function will display the number of rows in a dataset.

```{r}
ncol(akweather)
```
The ncol function will display the number of column in a dataset.

**For the Weather dataset you chose, what levels of measurement are included?**

Within the dataset, the Tmax and Tmin are both ratio variables, which means that in addition to the measurement scale meeting the requirements of an interval variable, the ratios of values along the scale should be meaningful.

**Create a few vectors from your dataset and perform each of the following operations:**
```{r}
tmax = akweather$TMAX
tmin = akweather$TMIN
```

**max(x)**
```{r}
max(tmax)
max(tmin)
```
**min(x)**
```{r}
min(tmax)
min(tmin)
```
**sum(x)**
```{r}
sum(tmax)
sum(tmin)
```
**mean(x)**
```{r}
mean(tmax)
mean(tmin)
```
**median(x)**
```{r}
median(tmax)
median(tmin)
```
**range(x)**
```{r}
range(tmax)
range(tmin)
```
**var(x)**
```{r}
var(tmax)
var(tmin)
```
**sd(x)**
```{r}
sd(tmax)
sd(tmin)
```

**Choose 4 of the operations above and provide an explanation of the result for each.**
```{r}
min(tmax)
min(tmin)
```
The min function pulls the lowest value from each of the vectors. Therefore, in this context, these are the coldest days in the dataset.

```{r}
max(tmax)
max(tmin)
```
The max function pulls the highest value from each of the vectors. Therefore, in this context, these are the warmest days in the dataset.

```{r}
median(tmax)
median(tmin)
```
The median function pulls the "middle" value in a result set. It divides a distribution in half, meaning that half of the observations lie abve the median and half lie below. (If there is an even number of observations, the median is the midpoint between the two middle observations.) Therefore, in this context, this is a great estimate for the average highest temperture, and the average coldest temperture per day.

```{r}
sum(tmax)
sum(tmin)
```
The sum is simply the aggregate of all the numerical values in a dataset. However, in this context, this function provides us little, to no value, since the cumulative highest temperture doesn't give us any new information without having another figure to compare it to. 
