---
title: 'DSC 520 Assignment 3.1: 2014 American Community Survey'
author: "Demond Love"
date: "6/24/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**The course professor will set up groups (2 students per group) for this assignment. Data for this assignment originated from the 2014 American Community Survey, which is conducted monthly by the Census Bureau and is designed to show how communities are changing. Through asking questions of a sample of the population, it produces national data on more than 35 categories of information, such as education, income, housing, and employment.**

**For this assignment, you will need to load and activate the ggplot2 package. You are encouraged to complete the assigned reading and DataCamp exercises before starting your work. Each of you should first produce your own version of the following deliverables then share with your team members. Team members will then collaborate on their approach and insights, refining their work before submitting for a grade. The following guidelines describe minimum deliverables needed for the assignment submission.**

```{r beginning, results="hide"}
library(readr)
library(ggplot2)
library(pastecs)
path = file.path('/Users/Love/Documents/DSC 520 Statistics for Data Science/Week 3 DSC 520/DSC 520 Week 3/acs-14-1yr-s0201.csv')
file = read_csv(path, col_names = TRUE)
```

**1. Document the elements in your data including the categories and data types.**

There are 8 categories in the file. Id, which is a character field, Id2, which is a integer field, Geography, which is a character field, PopGroupID, which is an integer field, POPGROUP.display-label, which is a character field, RacesReported, which is an integer field, HSDegree, which is a double field, and BachDegree, which is a double field. 

**2. Report the output from at least the following functions: str(); nrow(); ncol()**
```{r}
str(file)
nrow(file)
ncol(file)
```

**3. Create a Histogram of the HSDegree variable using the ggplot2 package. Set a bin size for the Histogram.**
```{r}
surveydata = ggplot(file, aes(file$HSDegree), title = 'High School Degrees')
```

**4. Include a Title and appropriate X/Y axis labels on your Histogram Plot.**
```{r}
surveydata + geom_histogram(binwidth = 0.8) + labs(x = 'Percentage with High School Degrees', y = 'Frequency')
```

**5. Answer the following questions based on the Histogram produced:**
**a. Based on what you see in this histogram, is the data distribution unimodal?**
Yes.

**b.  Is it approximately bell-shaped?**
No.

**c.  Is it approximately normal?**
No.

**d.  If not normal, is the distribution skewed? If so, in which direction? Provide specifics in your responses.**
Yes, it is a negatively/ left-skewed distribution.

**e.   Include a normal curve to the Histogram that you plotted.**
```{r}
surveydata = ggplot(file, aes(file$HSDegree), title = 'High School Degrees') + stat_function(fun = dnorm)
surveydata + geom_histogram(binwidth = 0.8) + labs(x = 'Percentage with High School Degree', y = 'Frequency')
```

**f.   Explain whether a normal distribution can accurately be used as a model for this data.**
No, a normal distribution cannot be used to accurately describe this distribution, because the HSDegree is negatively skewed.

**6. Create a Probability Plot of the HSDegree variable.**
```{r}
qplot(sample = HSDegree, data = file)
```

**7. Answer the following questions based on the Probability Plot:**
**a. Based on what you see in this probability plot, is the distribution approximately normal? Explain how you know.**
No, you can tell because if the values fall on the diagonal of the plot then the variable is normally distributed, but deviations from the diagonal show deviations from normality.

**b. If not normal, is the distribution skewed? If so, in which direction? Explain how you know.**
Yes, this distribution is negatively skewed, and you can tell because the distribution is deviating from the diagonal, and is bowing out above it.

**8. Now that you have looked at this data visually for normality, you will now quantify normality with numbers using the stat.desc() function. Include a screen capture of the results produced.**
```{r}
round(stat.desc(file$HSDegree, basic = FALSE, norm = TRUE), digits = 3)
```

**9. In several sentences provide an explanation of the result produced for skew, kurtosis, and z-scores. In addition, explain how a change in the sample size may change your explanation?**
The values of skew and kutosis should be zero in a normal distribution. However, skewness is a -1.675, which indicates a significant pile-up on the right. And a kutosis of 4.353 indicates a pointy and heavy-tailed distribution. Lastly, the z-scores show that these are statistically significant, above the 0.001 confidence level with a skew.2SE score of -4.030 and hurt.2SE score of 5.274.

**10.3. Create a Histogram of the HSDegree variable using the ggplot2 package. Set a bin size for the Histogram.**
```{r}
surveydata2 = ggplot(file, aes(file$BachDegree), title = 'Bachelors Degrees')
```


**10.4 Include a Title and appropriate X/Y axis labels on your Histogram Plot.**
```{r}
surveydata2 + geom_histogram(binwidth = 0.8) + labs(x = 'Percentage with Bachelors Degrees', y = 'Frequency')
```


**10.5. Answer the following questions based on the Histogram produced:**
**a. Based on what you see in this histogram, is the data distribution unimodal?**
Yes.

**b.  Is it approximately bell-shaped?**
Yes.

**c.  Is it approximately normal?**
Yes.

**d.  If not normal, is the distribution skewed? If so, in which direction? Provide specifics in your responses.**
No, it is not a skewed distribution.

**e.   Include a normal curve to the Histogram that you plotted.**
```{r}
surveydata2 = ggplot(file, aes(file$BachDegree), title = 'Bachelors Degrees') + stat_function(fun = dnorm) + geom_histogram(binwidth = 0.8) + geom_histogram(binwidth = 0.8) + labs(x = 'Percentage with Bachelors Degrees', y = 'Frequency')
surveydata2
```

**f.   Explain whether a normal distribution can accurately be used as a model for this data.**
Yes, a normal distribution can be used to accurately describe this distribution, because the BachDegree is nicely symmetrical and doesn't seem too pointy or flat.

**10.6. Create a Probability Plot of the HSDegree variable.**
```{r}
qplot(sample = BachDegree, data = file)
```

**10.7. Answer the following questions based on the Probability Plot:**
**a. Based on what you see in this probability plot, is the distribution approximately normal? Explain how you know.**
Yes, you can tell because the data has the same distribution as that of a normal distribution, and you can see a lovely straight diagonal line.

**b. If not normal, is the distribution skewed? If so, in which direction? Explain how you know.**
This is a normal distribution, without statistically significant skew.

**10.8. Now that you have looked at this data visually for normality, you will now quantify normality with numbers using the stat.desc() function. Include a screen capture of the results produced.**
```{r}
round(stat.desc(file$BachDegree, basic = FALSE, norm = TRUE), digits = 3)
```

**10.9. In several sentences provide an explanation of the result produced for skew, kurtosis, and z-scores. In addition, explain how a change in the sample size may change your explanation?**
The values of skew and kutosis should be zero in a normal distribution. Our skewness value is 0.328 and our kurtosis value is -0.277 If skew.2SE or kurt.2SE are greater than 1 (ignoring the plus or minus sign) then you have significant skew/kurtosis. However, in our distribution we skew.2SE and kurt.2SE scores of 0.790 and -0.336 respectively. Therefore, our z-scores do not dictate that our skewness or kurtosis are statistically significant.
